---
title: "Exploring the link between two variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring_the_link_between_two_variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE,message=FALSE
)
```

```{r setup}
library(dplyr)
library(questionable)
library(randomForest)
data(icecream)
```


# Plotting the relationship between two variables

When trying to explore the link between two variables X and Y, you can use function `biplot()`. It adapts to the quantitative or categorical nature of variables X and Y.

## Two categorical variables

```{r,fig.width=5, fig.height=4, warning=FALSE, message=TRUE}
biplot(icecream,"creme_glacee_vanille","creme_glacee_chocolat")
test_pval(icecream,"creme_glacee_vanille","creme_glacee_chocolat")
```

In this case, a chi-squared test is performed. The p-value is calculated and displayed in the plot's title. The p-value is also accessible via the function `test_pval()`.

## One quantitative, one categorical variable


```{r, fig.width=5, fig.height=4, warning=FALSE, message=FALSE}
biplot(icecream,"age","creme_glacee_chocolat")
test_pval(icecream,"age","creme_glacee_chocolat")
```

In this case, a ANOVA is performed.

## Two quantitative variables


```{r, fig.width=5, fig.height=4, warning=FALSE, message=FALSE}
biplot(icecream,"age","taille")
```

In the case of a non-linear regression (as displayed above) no test is carried out.

Running a linear regression (method="lm" for **L**inear **M**odel) allows a test to be performed.

```{r, fig.width=5, fig.height=4, warning=FALSE, message=FALSE}
biplot(icecream,"age","taille",method="lm")
test_pval(icecream,"age","taille")
```

# General overview of the p-values

For a general overview of all variables' inter-relationships, you can use the function `pval_graph()`. This function will calculate the p-values for all pairs of variables in the dataset. The resulting graph will display the p-values in a heatmap format, allowing for easy identification of significant relationships.

```{r, fig.width=10, fig.height=6}
data(icecream)
pval_graph(icecream %>% dplyr::select(-comment,-id), var_reorder=TRUE)
```

