---
title: "Intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE,
  message=FALSE
)
```

```{r setup}
library(questionable)
library(dplyr)
```

# Objectives of the package 

This package aims at making the analysis of questionnaires easier. It provides functions to

- **clean and transform data** in order to make it useable for analysis (in particular, running random forests)
- **run random forests** in order to **identify which variables are the most important for predicting a given outcome**
- **test and visualize the link between two variables**

These three aspects are dealt with in three different vignettes.

The package comes with a built-in dataset (of purely fantasy origin) called `icecream` which helps examplify the functions and general purpose of the package. This hence questionable data contains the **results of a questionnaire about ice cream consumption**. The dataset is used in the examples below.

```{r}
data(icecream)
```

This dataset contains `nrow(icecream)` rows corresponding to the number of respondants and `ncol(icecream)` columns. These columns (or variables) are of varying types and are described below:

```{r}
icecream %>% 
  dplyr::glimpse()
```
