---
title: "Running random forests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running_random_forests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE
)
```

```{r setup}
library(dplyr)
library(questionable)
library(randomForest)
data(icecream)
```

# First exploration

Now, let's explore how **some ice-cream consumption characteristics relate to other characteristics**.

We prepare the data for running a random forest:

```{r}
icecream_for_rf=icecream %>%
          select(-id,-comment,-creme_glacee_caramel) %>%
          mutate_all(optimize_for_rf)
```

Let's start with exploring what could explain the **regularity of eating ice-cream**. Here is the distribution of the variable:

```{r}
icecream_for_rf %>% 
  group_by(regular_eating) %>% 
  tally()
```

# Running a "wild guess" random forest

Imagine you have no idea whatsoever about possible explaining factors. Let's just run a random forest with all possible predictors.

```{r, fig.height=5, fig.width=7}
rf=run_rf(icecream_for_rf,
          response="regular_eating",
          plot=TRUE,
          min_importance=0)
```

This displays the **relative importance of the predictors** in predicting the response variable.

Now, this importance is only interpretable insofar as the model performs well. Let's check the performance of the model:

```{r}
rf$performance
```

Here, we're dealing with a categorical response. The overall performance corresponds to the model predicting correctly the "regular_eating" response in about `r round(rf$performance,2)*100` % of the cases. This should be mitigated by the fact that the response is not balanced. We can check this by looking at the confusion matrix:

```{r}
rf$rf$confusion
```
The "Yes"s and "No"s are actually predicted with quite low performance due to frequent confusion with the "Non renseigné"s.




# Understanding the predictor's effect

Now, let's for instance consider the effect of age on whether or not people regularly eat ice cream.

```{r, fig.height=4, fig.width=5}
partial_dependence_plot(rf=rf,
                        predictor="age",
                        probabilities=TRUE)
```

The plot above shows the **normalized logit scores** for the **probability of being a regular eater of ice cream**. The higher the score, the more likely it is that a person is a regular eater of ice cream. Here in particular, we can see that younger people are more likely to assert themselves as regular eaters of ice cream.

# On collinearity

According to the random forest importance assessment of predictors' importance, **both age and size of the respondant are important predictors**. Well. As an omniscient creator of the data, I can tell you that, yes, indeed answers regarding regular eating of icecream were made dependent on age. On the other hand, I did not make them dependent on the size of the respondant, which would be (from my non-expert point of view on ice-cream consumption anyway) an odd choice.

On the other hand, I did make the answer dependent on gender, and I did make size dependent on age and gender. A not-too-wild guess is that size somehow **integrates the effect of more relevant predictors**, making it appear relatively more important than it actually is, and making these other, more meaningful predictors appear less important than they actually are.

This hence illustrate that although random forests are a great tool for exploratory data analysis, they are not a substitute for a good understanding of the data and the domain (and certainly not a free card to include probably irrelevant descriptors in the data).

So, let's run a random forest excluding variable `taille` first, which will probably be a safer basis for later interpretations.

```{r, fig.height=5, fig.width=7}
rf=run_rf(icecream_for_rf %>% 
            dplyr::select(-taille),
          response="regular_eating",
          plot=TRUE,
          min_importance=0)
```

You can see that the relative importance of gender appears higher now that we have removed size from the mix.
